Technique for \earlywarning\ detection of compact binary coalescence and its implications for multi-messenger astronomy

### 2011-06-11 introduction.tex
Moved discussion of Figure 1 to main text
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1017 ccdd05d5-56c1-461a-bf98-85f6876523b4

We show that by examining the signal to noise ratio (\SNR{}) stream for
threshold crossings before the \GW{} signal leaves the detection band, it is
possible to trade some \SNR{} and sky localization
accuracy for negative latency.  Figure~\ref{fig:earlywarning} shows
projected early trigger rates for \textsc{ns}--\textsc{ns} binaries in Advanced
\LIGO\ assuming event rates predicted in \citet{Abadie:2010p10836} and anticipated
detector sensitivity described in \citet{ALIGONoise}.  The
most likely estimates indicate that $\sim$10 sources will produce an \SNR\
above 8 in the detector 10 seconds or more prior to the merger during one
year of observation, though the uncertainty spans orders of magnitude.
The gray bands give the 5 to 95\% confidence interval as described
in \citet{Abadie:2010p10836}.
%
\begin{figure}[h]
\plotone{figures/snr_in_time.pdf}
\caption{\label{fig:earlywarning} Expected number of \textsc{ns}--\textsc{ns}
sources that will be detectable $t$ seconds before coalescence.  The heavy
solid line is the most likely yearly rate estimate $\dot N_{\mathrm{re}}$
during Advanced \LIGO.  The shaded region represents the confidence interval
$[\dot N_{\mathrm{low}}, \dot N_{\mathrm{high}}]$ described in
\citep{Abadie:2010p10836}.  With an advanced detector in the `zero detuning,
high power' configuration \citep{ALIGONoise} and an \SNR\ threshold of 8,
we will observe a total of $\dot N_{\mathrm{re}} = 40$ events~yr$^{-1}$, but
$\sim$10~yr$^{-1}$ will be detectable within 10~s of merger and
$\sim$1~yr$^{-1}$ within 100~s.}
\end{figure}

>>>>

We show that by examining the signal to noise ratio (\SNR{}) stream for
threshold crossings before the \GW{} signal leaves the detection band, it is
possible to trade some \SNR{} and sky localization
accuracy for negative latency.  Figure~\ref{fig:earlywarning} shows
projected early detection rates for \textsc{ns}--\textsc{ns} binaries in Advanced
\LIGO\ assuming event rates predicted in \citet{Abadie:2010p10836} and anticipated
detector sensitivity for the `zero detuning, high power' configuration described in 
\citet{ALIGONoise}.  The most realistic estimates indicate that at an \SNR\ threshold
of 8, we will observe a total of 40~events~yr$^{-1}$, but $\sim$10~yr$^{-1}$ will be
detectable within 10~s of merger and $\sim$1~yr$^{-1}$ will be detectable within
100~s of merger.  These rates are uncertain by an order of magnitude.
%
\begin{figure}[h]
\plotone{figures/snr_in_time.pdf}
\caption{\label{fig:earlywarning}Expected number of \textsc{ns}--\textsc{ns}
sources that will be detectable by Advanced \LIGO\ $t$ seconds before coalescence.
The heavy solid line is the most realistic yearly rate estimate.  The shaded
region represents the 5 to 95\% confidence interval arising from uncertainty
in predicted event rates \citep{Abadie:2010p10836}.}
\end{figure}
